Linear analysis of the vertical shear instability in astrophysical disks

### 2015-01-24 application.tex
regular update, code updates

Fig. \ref{mmsn_bcrit_bcool} suggests that the fundamental isothermal
VSI can operate in the outer parts of the MMSN at radii on the order
of tens of AU. However, the thermal relaxation time $\beta$ is not
much shorter than $\beta_\mathrm{crit}$, suggesting the VSI may be
weak.

@@

\subsection{Numerical example}
Consider perturbations with $\khat=10$ in the 
the MMSN at $r=30\mathrm{AU}$, with $\gamma=1.4$ and
$\mu=2$. Then we find
the dimensionless thermal relaxation time, using
Eq. \ref{tc_newton_cool}, is
\begin{align*}
  \beta \simeq 0.046. 
\end{align*}
At $r=30\mathrm{AU}$, the disk aspect ratio is $\epsilon \simeq
0.058$ which implies, using Eq. \ref{iso_vsi_cond}, a critical 
relaxation timescale of  
\begin{align*}
  \beta_\mathrm{crit} \simeq 0.062. 
\end{align*}
Thus, the disk supports the fundamental isothermal VSI, but only
marginally. Using the above value of $\beta$, MMSN disk parameters
and $\khat=10$ in our linear code, we find a growth rate of
\begin{align*}
  \nu = 2.7\times10^{-3}\Omega_k.
\end{align*}
This corresponds to a characteristic growth time of $t_g\simeq 60$
orbits.

 %only find one unstable mode -> close to marginal

 %larger kx

>>>>

Fig. \ref{mmsn_bcrit_bcool} suggests that the fundamental isothermal
VSI can operate in the outer parts of the MMSN at radii on the order
of tens of AU. However, the thermal relaxation time $\beta$ is not
much shorter than $\beta_\mathrm{crit}$, so the VSI may be
weak.

@@

\subsection{Numerical example}
Consider  the MMSN at $r=30\mathrm{AU}$, with $\gamma=1.4$ and
$\mu=2$. Eq. \ref{mmsn_epsilon} gives $\epsilon \simeq
0.058$ which implies, using Eq. \ref{bcrit_mmsn}, a critical 
relaxation timescale of  
\begin{align*}
  \beta_\mathrm{crit} \simeq 0.062. 
\end{align*} 
To estimate a thermal relaxation timescale, we set $\khat=20$ 
in Eq. \ref{beta_mmsn} to obtain 
\begin{align*}
  \beta \simeq 0.042
\end{align*}
at $r=30\mathrm{AU}$. This indicates the disk supports the fundamental
isothermal VSI, but only marginally. Using the above value of $\beta$,
MMSN disk parameters and $\khat=20$ in our linear code, we find a
growth rate of 
\begin{align*}
  \nu = 3.3\times10^{-3}\Omega_k.
\end{align*}
This corresponds to a characteristic growth time of $t_g\sim 50$
orbits. In fact, this mode was found to be the most unstable for
$\khat=20$. 
% and
% physically acceptable -> other modes
% are crazy

We also solved the linear problem with varying $\khat$  
and $\beta$ determined by Eq. \ref{beta_mmsn}. We find the most
unstable wavenumber occurs at $\khat\simeq60$ with $\nu\simeq
3.6\times10^{-3}\Omega_k$, i.e. only slightly faster than that above. 
However, the eigenfunctions for large $\khat$ were found to have large
amplitude oscillations near the vertical boundaries.

 %only find one unstable mode -> close to marginal 
 %larger kx
